Query Optimization for Dynamic Imputation

### 2017-10-19 paper/algorithm.tex
fix: change lang in table to avoid overflow into margin
paper: bold all figure labels

\begin{table}
  \centering
  \caption{\cameraready{Notation used in Section~\ref{sec:algorithm}.}}
\cameraready{
  \begin{tabular}{ll}
    \toprule
    Operator & Description \\
    \midrule
    $\textsc{Attr}(\cdot)$ & Attributes used in a predicate. \\
    $\textsc{Time}(\cdot)$ & Estimated time to run a plan. \\
    $\textsc{Penalty}(\cdot)$ & Imputation quality penalty of a plan. \\
    $\textsc{Card}(\cdot)$ & Estimated cardinality of a plan. \\
    $\widehat{Var}(\cdot)$ & Estimated variance of a column. \\
    $\textsc{Dirty}(\cdot)/\textsc{Clean}(\cdot)$ & Dirty/clean attributes in a plan. \\
    $Q[\cdot]$ & Plan cache access operator. \\
    $\cdot \triangleleft \cdot$ & Plan cache insertion operator. \\
    \bottomrule
  \end{tabular}
}
\end{table}

@@

\begin{figure}
  \begin{minipage}[c]{0.5\columnwidth}
    \Tree [.$\pi$ [.$g$ [.$\bowtie$ [.$\bowtie$ [.$\sigma$ $t_1$ ] [.$\sigma$ $t_2$ ] ] [.$\sigma$ $t_3$ ] ] ] ]
  \end{minipage}\begin{minipage}[c]{0.5\columnwidth}
    \caption{
       Query plan schematic for the type of traditional plans explored (absent imputation operators).
    } \label{fig:query-schematic}
  \end{minipage}
\end{figure}

@@

Drop operators (\Crefrange{lst:drop-start}{lst:drop-end}) reduce the cardinality of their input relation by removing tuples which contain missing values.
For a drop $\delta_C$, for each column $c \in C$ we set its missing value count to zero, because all of these tuples will be dropped.
This can result in a set of histograms which have different cardinalities if some columns have more missing values than others.
To account for the discrepancy, we rescale all the histograms without changing the shape of their distributions to match the lowest cardinality column.

Impute operators (\Crefrange{lst:impute-start}{lst:impute-end}) maintain the cardinality of their inputs, but they also fill in missing values.
For an impute $\mu_C$, we set the missing value count to zero, and rescale each histogram so that its new cardinality, with no missing values, is the same as its old cardinality.
We assume that the imputed values will not change the shape of the distribution of values in each column.

@@

\begin{figure}
  \cameraready{
  \begin{align*}
    \llbracket Q[T] \rrbracket &= \begin{cases}
      P & (T,\ P) \in Q \\
      \emptyset & \text{otherwise} \\
    \end{cases} \\
    \llbracket Q[T] \lhd p \rrbracket &= \begin{aligned} Q \cup \{(T,\ P) \} \end{aligned} \text{ where }\\
    P &= \left\{q \in S ~|~ \begin{aligned}
              \textsc{Dirty}(q) \neq \textsc{Dirty}(p) \lor
              \not \exists q' \in S:\ \doms{q'}{q}
              \end{aligned}
            \right\} \\
    S &= \llbracket Q[T] \rrbracket \cup \{p\}
  \end{align*}
  \caption{\cameraready{Semantics of the plan cache access and insertion syntax. We use $\llbracket\ \rrbracket$ to mean evaluation. $Q$ is a plan cache, $T$ is a set of tables, and $p$ is a query plan. The cache guarantees that plans for the same set of tables either have distinct dirty sets or are Pareto optimal.}}
  \label{fig:semantics-plan-cache}
  }
\end{figure}

>>>>

\begin{table}
  \centering
  \captionsetup{labelfont=bf}
  \caption{\cameraready{Notation used in Section~\ref{sec:algorithm}.}}
  \cameraready{
  \begin{tabular}{ll}
    \toprule
    Operator & Description \\
    \midrule
    $\textsc{Attr}(\cdot)$ & Attributes used in a predicate. \\
    $\textsc{Time}(\cdot)$ & Estimated time to run a plan. \\
    $\textsc{Penalty}(\cdot)$ & Impute quality penalty of a plan. \\
    $\textsc{Card}(\cdot)$ & Estimated cardinality of a plan. \\
    $\widehat{Var}(\cdot)$ & Estimated variance of a column. \\
    $\textsc{Dirty}(\cdot)/\textsc{Clean}(\cdot)$ & Dirty/clean attributes in a plan. \\
    $Q[\cdot]$ & Plan cache access operator. \\
    $\cdot \triangleleft \cdot$ & Plan cache insertion operator. \\
    \bottomrule
  \end{tabular}
}
\end{table}

@@

\begin{figure}
  \captionsetup{labelfont=bf}
  \begin{minipage}[c]{0.5\columnwidth}
    \Tree [.$\pi$ [.$g$ [.$\bowtie$ [.$\bowtie$ [.$\sigma$ $t_1$ ] [.$\sigma$ $t_2$ ] ] [.$\sigma$ $t_3$ ] ] ] ]
  \end{minipage}\begin{minipage}[c]{0.5\columnwidth}
    \caption{
       Query plan schematic for the type of traditional plans explored (absent imputation operators).
    } \label{fig:query-schematic}
  \end{minipage}
\end{figure}

@@

\textit{Drop} operators (\Crefrange{lst:drop-start}{lst:drop-end}) reduce the cardinality of their input relation by removing tuples which contain missing values.
For a \textit{Drop} $\delta_C$, for each column $c \in C$ we set its missing value count to zero, because all of these tuples will be dropped.
This can result in a set of histograms which have different cardinalities if some columns have more missing values than others.
To account for the discrepancy, we rescale all the histograms without changing the shape of their distributions to match the lowest cardinality column.

\textit{Impute} operators (\Crefrange{lst:impute-start}{lst:impute-end}) maintain the cardinality of their inputs, but they also fill in missing values.
For an \textit{Impute} $\mu_C$, we set the missing value count to zero, and rescale each histogram so that its new cardinality, with no missing values, is the same as its old cardinality.
We assume that the imputed values will not change the shape of the distribution of values in each column.

@@

\begin{figure}
  \captionsetup{labelfont=bf}
  \cameraready{
  \begin{align*}
    \llbracket Q[T] \rrbracket &= \begin{cases}
      P & (T,\ P) \in Q \\
      \emptyset & \text{otherwise} \\
    \end{cases} \\
    \llbracket Q[T] \lhd p \rrbracket &= \begin{aligned} Q \cup \{(T,\ P) \} \end{aligned} \text{ where }\\
    P &= \left\{q \in S ~|~ \begin{aligned}
              \textsc{Dirty}(q) \neq \textsc{Dirty}(p) \lor
              \not \exists q' \in S:\ \doms{q'}{q}
              \end{aligned}
            \right\} \\
    S &= \llbracket Q[T] \rrbracket \cup \{p\}
  \end{align*}
  \caption{\cameraready{Semantics of the plan cache access and insertion syntax. We use $\llbracket\ \rrbracket$ to mean evaluation. $Q$ is a plan cache, $T$ is a set of tables, and $p$ is a query plan. The cache guarantees that plans for the same set of tables either have distinct dirty sets or are Pareto optimal.}}
  \label{fig:semantics-plan-cache}
  }
\end{figure}
